\begin{thebibliography}{99}
  
\bibitem{SixTrack_user_manual}
  F.~Schmidt, ``SixTrack: Version 4.7.16, Single Particle
  Tracking Code Treating Transverse Motion with Synchrotron
  Oscillations in a Symplectic Manner, User's Reference Manua'',
  CERN/SL/94--56 (AP), CERN, Geneva, Switzerland (2017),
  \url{http://sixtrack.web.cern.ch/SixTrack/docs/user_full/manual.php}
  
\bibitem{SixPub}
    G.~Ripken and F.~Schmidt,
    ``A symplectic six-dimensional thin-lens formalism for tracking'',
    CERN, Geneva, Switzerland, Rep.~CERN/SL/95--12(AP), 1995.

\bibitem{sixtrackWeb} SixTrack, \url{http://sixtrack.web.cern.ch/SixTrack}

\bibitem{SixDesk_original}
  M.~Hayes and F.~Schmidt, ``Run Environment for SixTrack'',
  LHC Project Note 300, CERN, Geneva, Switzerland (2002),
  \url{https://cds.cern.ch/record/691785/files/project-note-300.pdf}

\bibitem{SixDesk_updated}
  E.~McIntosh and R.~De Maria, ``The SixDesk Run Environment for SixTrack'',
  CERN-ATS-TE-2012-089 TECH, CERN, Geneva, Switzerland (2002),
  \url{https://github.com/SixTrack/SixDesk/blob/master/sixjobs/doc/sixdesk_env.pdf}

\bibitem{DynApeStudiesGiovannozzi2015}
  M.~Giovannozzi \emph{et al.},
  ``Dynamic Aperture Studies for the LHC Hight Luminosity Lattice'',
  in \emph{Proc.~6$^{\textrm{th}}$ Int.~Particle Accelerator Conf.~(IPAC'15)},
  Richmond, VA, USA, May 2015, paper MOPMN003, 
  pp.~705--709.
  
\bibitem{NomLHCdesignRepoV1}
  O.~Br\"{u}ning \emph{et al.} (eds), ``LHC design report'', Vol.~I,
  CERN, Geneva, Switzerland, Rep.~CERN-2004-003-V-1, 2004.
  
\bibitem{HLLHC_book}
  O.~Br\"{u}ning, L.~Rossi (eds.),
  ``The High Luminosity Large Hadron Collider'',
  World Scientific Press, 2015, ISBN 978-981-4675-46-8.
  
\bibitem{HLLHCtechDesRepo}
  G.~Apollinari, I.~Bejar Alonso, O.~Br\"{u}ning, M.~Lamont, L.~Rossi (eds.),
  ``High Luminosity Large Hadron Collider (HL-LHC) Technical Design Report
  V.01'', CERN, Geneva, Switzerland, EDMS n.~1723851 v.0.71,
  \url{https://edms.cern.ch/ui/file/1723851/0.71/HL_TDR_V07.0.2016.10.05.Version15.2h.pdf}
  
\bibitem{SixTrackForCollimation}
  G.~Robert-Demolaize, R.~Assmann, S.~Redaelli, F.~Schmidt,
  ``A New Version of SixTrack with Collimation and Aperture Interface'',
  in \emph{Proc.~Particle Accelerator Conf.~(PAC'05)}, 
  Knoxville, TN, USA, 2005, paper FPAT081,
  pp.~4084--4087.
    
\bibitem{LHCCollSys}
  R.W.~Assmann \emph{et al.},
  ``The Final Collimation System for the LHC'',
  in \emph{Proc.~10$^{\textrm{th}}$ European Particle Accelerator Conf.~(EPAC'06)},
  in \emph{Proc.~EPAC'06},
  Edinburgh, Scotland, UK, Jun 2006, paper TUODFI01,
  pp.~986--988.
  
\bibitem{RHICcollSys}
  G.~Robert-Demolaize and A.~Drees, ``Simulations of collimation
  losses at RHIC'', in \emph{Proc.~Tracking for Collimation Workshop},
  CERN, Geneva, Switzerland, Oct 2015, unpublished.

\bibitem{HLLHCTrackWS}
  S.~Redaelli (ed.), ``Proceedings of the tracking for collimation
  workshop'', CERN, Geneva, Switzerland, Oct 2015, unpublished.

\bibitem{Amereghe6TColl}
  A.~Mereghetti \emph{et al.},
  ``SixTrack for Cleaning Studies: 2017 Updates'',
  in \emph{Proc.~8$^{\textrm{th}}$ Int.~Particle Accelerator Conf.~(IPAC'17)},
  Copenhagen, DK, May 2017, paper THPAB046, 
  pp.~3811--3814.
  
\bibitem{Boinc} ``Berkeley Open Infrastructure for Network Computing'',
  \url{http://boinc.berkeley.edu}
  
  \addcontentsline{toc}{chapter}{Bibliography}
\end{thebibliography}

